National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: tools to support monitoring decisions

### 2023-07-24 analysis/paper/paper.Rmd
in progress - bias term.

The composition survey bias correction term $c$ depends on the apparent number of adult females per collared animal $w$, the ratio of young bulls to adult females $q$, the probability of misidentifying young bulls as adult females and vice versa $u$, and the probability of missing a calf $z$ (Appendix X):
$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$$
In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Gaussian distribution (Appendix X). Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $c$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.

@@

```{r biasDistribution, fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Gaussian distribution (red). In these examples, the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2. We assume that the apparent number of adult females per collared animal $w$ is known.", cache=TRUE}

@@

#get mean and sd as fn of w
cs <- cr %>%
  group_by(w) %>%
  summarise(mean = mean(c), sd = sd(c))

cr = merge(cr,cs)
cr$ca= rnorm(nrow(cr),cr$mean,cr$sd)

>>>>

The composition survey bias correction term $c$ depends on the apparent number of adult females per collared animal $w$, the ratio of young bulls to adult females $q$, the probability of misidentifying young bulls as adult females and vice versa $u$, and the probability of missing a calf $z$ (Appendix X):
$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$$
In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Log-normal distribution (Appendix X). Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $\log{c}$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.

@@

```{r biasDistribution, fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Log-normal distribution (red). In these examples, the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2. We assume that the apparent number of adult females per collared animal $w$ is known.", cache=TRUE}

@@

#get mean and sd as fn of w
cs <- cr %>%
  group_by(w) %>%
  summarise(mean = mean(log(c)), sd = sd(log(c)))

cr = merge(cr,cs)
cr$ca= exp(rnorm(nrow(cr),cr$mean,cr$sd))
